---
title: "writeup"
output: html_document
---

The first step is to declare our libraries and import the data.
Note we will only be importing the training set at this time.


```{r}
data<-read.csv("pml-training.csv")
```
Note that we get a table with 160 variables. classe is our Y variable, and X is just index.
A lot of the columns have missing data in them. Im going to get rid of any columns with NA.

remColumns<-grep("NA",data[1,])
data2<-data[,-remColumns]
data3<-data2[,(data2[1,]!="")]

Ill also get rid of any columns where the data is blank
data3<-data2[,(data2[1,]!="")]

At this point we have 60 variables, including our Y variable, and index variable.
Im going to get rid of the index and the username variables, since those might skew our predictions.
We dont want to have a model that says "If its carlitos, then he is not doing the exercise right"
We want to look at the individual exercise and know if its bad or good based on the data.

data4<-data3[!names(data3) %in% c("X","user_name")]

Now we have 58 features, and 19622 rows where each row is an independent example.
We will now take the data and split it into testing and training. 
library(caret)
set.seed(4582)
inTrain<-createDataPartition(y=data4$classe,p=.75,list=FALSE)
training<-data4[inTrain,]
testing<-data4[-inTrain,]
model1<-train(classe~.,data=training,method="rf")
boostFit <- train(classe ~ ., method = "gbm", data = training, verbose = F, trControl = fitControl)

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
